---
title: 'LDA VS QDA'
author: 'Mohd Javed'
date: "`r Sys.Date()`"
output:
  pdf_document: default
  word_document: default
---

# Part 1


```{r}
library(Ecdat)
library(MASS)
```

### Linear Discreminant Analysis


```{r}

# training lda model
model_1<-lda(choice~.,data=Mode)

model_1
```

### Quadratic discriminant analysis

```{r}
set.seed(123)
# training qda model
model_2<-qda(choice~.,data=Mode)
model_2



```

### Confusion Matrix for Both model

```{r}


#Confusion matrix for lda model

table(predict(model_1,type="class")$class,Mode$choice)

#Confusion matrix for qda model

table(predict(model_2,type="class")$class,Mode$choice)



```


## Sample error rate

```{r}
# Incorrectly Classified Instances for lda model
(1-mean(predict(model_1,type="class")$class==Mode$choice))*100

# Incorrectly Classified Instances for qda model

(1-mean(predict(model_2,type="class")$class==Mode$choice))*100

```
### Quadratic Discriminent Anlysis performing better than Linear Discriminent Anlysis beacuse it has lower sample classification error rate.

# Part 2

```{r}
library(ISLR2)

```

```{r}
#Creating copy of dataframe
Default_LM<-Default
#Converting factor to numeric
Default_LM$default <- ifelse (Default_LM$default == 'Yes', 1, 0)
#training the linear probability model
mod_lm<-lm(default~.,data=Default_LM)
summary(mod_lm)

Default_LM$Predict_lm <- ifelse(mod_lm$fitted.values > 0.5,1,0)
table(Default_LM$Predict_lm,Default_LM$default)

mean(Default_LM$Predict_lm==Default_LM$default)

```


```{r}
Default_LM$Predict_lm=NULL
mod_glm<-glm(default~.,data=Default_LM)
summary(mod_glm)

predictions_glm<-predict(mod_glm,Default_LM)
Default_LM$Predict_glm <- ifelse(predictions_glm > 0.5,1,0)
table(Default_LM$Predict_glm,Default_LM$default)
mean(Default_LM$Predict_glm==Default_LM$default)

```
## Both Models has the same accuracy/sample error rate. Both models are performing same toward classification of response variable 'default'.









```{r}
library(boot)
summary(Strike)

bootFunc <- function(data, i){
df <- data[i, ]
(1/nrow(df))*log(df$duration)
}

b <- boot(Strike, bootFunc, R = 62)
 
print(b)
 boot.ci(b, index = 1)


```
### From the output we can see that the 95% bootstrapped confidence interval for the true R-squared values is (.0000, .0599).


```{r}
### Partitioning the dataframe
set.seed(123)

library(caret)
sample_def <- createDataPartition(Default$default, p = 0.8, list = FALSE)
 
training_dataset  <- Default[sample_def, ]

validation_dataset <- Default[-sample_def, ]

```

# Default Dataset

### Linear Discreminant Analysis


```{r}
mod_def_lda<-lda(default~.,data=training_dataset)
mod_def_lda
```

### Quadratic discriminant analysis

```{r}
mod_def_qda<-qda(default~.,data=training_dataset)
mod_def_qda
```

### Confusion Matrix for Both model

```{r}
predictions_q<-predict(mod_def_qda,validation_dataset)
table(predictions_q$class,validation_dataset$default)
predictions_l<-predict(mod_def_lda,validation_dataset)
table(predictions_l$class,validation_dataset$default)


```
### Sample error rate

```{r}
1-mean(predictions_q$class==validation_dataset$default)
1-mean(predictions_l$class==validation_dataset$default)
```

### Quadratic Discriminent Anlysis performing better than Linear Discriminent Anlysis beacuse it has lower sample classification error rate.




# Participation Dataset

### Linear Discreminant Analysis



```{r}
### Partitioning the dataframe
set.seed(123)

sample_def <- createDataPartition(Participation$lfp, p = 0.8, list = FALSE)
 
training_dataset  <- Participation[sample_def, ]

validation_dataset <- Participation[-sample_def, ]

mod_par_lda<-lda(lfp~.,data=training_dataset)
mod_par_lda
```

### Quadratic discriminant analysis

```{r}
mod_par_qda<-qda(lfp~.,data=training_dataset)
mod_par_qda
```

### Confusion Matrix for Both model

```{r}
predictions_q<-predict(mod_par_qda,validation_dataset)
table(predictions_q$class,validation_dataset$lfp)
predictions_l<-predict(mod_par_lda,validation_dataset)
table(predictions_l$class,validation_dataset$lfp)
```
### Sample error rate

```{r}
1-mean(predictions_q$class==validation_dataset$lfp)
1-mean(predictions_l$class==validation_dataset$lfp)
```

### Quadratic Discriminent Anlysis performing better than Linear Discriminent Anlysis beacuse it has lower sample classification error rate


# Train Dataset

### Linear Discreminant Analysis


```{r}
### Partitioning the dataframe
set.seed(123)

sample_def <- createDataPartition(Train$choice, p = 0.8, list = FALSE)
 
training_dataset  <- Train[sample_def, ]

validation_dataset <- Train[-sample_def, ]
# training model
mod_trn_lda<-lda(choice~.,data=training_dataset)
mod_trn_lda
```

### Quadratic discriminant analysis

```{r}
# training model
mod_trn_qda<-qda(choice~.,data=training_dataset)
mod_trn_qda
```

### Confusion Matrix for Both model

```{r}
# conf.matrix for QDA
predictions_q<-predict(mod_trn_qda,validation_dataset)
table(predictions_q$class,validation_dataset$choice)
# conf.matrix for LDA
predictions_l<-predict(mod_trn_lda,validation_dataset)
table(predictions_l$class,validation_dataset$choice)
```
### Sample error rate

```{r}

1-mean(predictions_q$class==validation_dataset$choice)
1-mean(predictions_l$class==validation_dataset$choice)
```
### Quadratic Discriminent Anlysis performing better than Linear Discriminent Anlysis beacuse it has lower sample classification error rate.




# Mroz Dataset

### Linear Discreminant Analysis


```{r}
### Partitioning the dataframe
set.seed(123)

sample_def <- createDataPartition(Mroz$city, p = 0.8, list = FALSE)
 
training_dataset  <- Mroz[sample_def, ]

validation_dataset <- Mroz[-sample_def, ]
mod_Mroz_lda<-lda(city~.,data=training_dataset)
mod_Mroz_lda
```

### Quadratic discriminant analysis

```{r}
mod_Mroz_qda<-qda(city~.,data=training_dataset)
mod_Mroz_qda
```

### Confusion Matrix for Both model

```{r}
# conf.matrix for QDA

predictions_q<-predict(mod_Mroz_qda,validation_dataset)
table(predictions_q$class,validation_dataset$city)
# conf.matrix for LDA

predictions_l<-predict(mod_Mroz_lda,validation_dataset)
table(predictions_l$class,validation_dataset$city)
```
### Sample error rate

```{r}
1-mean(predictions_q$class==validation_dataset$city)
1-mean(predictions_l$class==validation_dataset$city)
```

### Linear Discriminent Anlysis performing better than Quadratic Discriminent Anlysis beacuse it has lower sample classification error rate.


