---
title: 'Socio economic Anaysis'
author: 'Mohd Javed'
date: "`r Sys.Date()`"
output: word_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,cache = TRUE,warning = FALSE,message = FALSE,dpi = 175,fig.width = 7.5,fig.height = 6)
```



### Importing required packages and Data 



```{r}
library(readr)
library(dplyr)
library(ggplot2)

homework <- read_csv("homework.csv")
View(homework)
```




## First look of the data


```{r}
skimr::skim(homework)
```



### Visualizing the Maths and English scores distribution by Socio-economic group



```{r}
ggplot(homework, aes(engach, fill=factor(ses))) + 
  geom_density(alpha=.5)

ggplot(homework, aes(mathsach, fill=factor(ses))) + 
  geom_density(alpha=.5)

```



## Descriptive measures for English Score by Socio-economic status



```{r}
group_by(homework, ses) %>%
  summarise(
    count = n(),
    mean = mean(engach, na.rm = TRUE),
    sd = sd(engach, na.rm = TRUE),
    median = median(engach, na.rm = TRUE),
    IQR = IQR(engach, na.rm = TRUE)
  )

ggplot(homework, aes(x = ses, y = engach, col = factor(ses))) + geom_boxplot(outlier.shape = NA) + geom_jitter(width = 0.2) + theme(legend.position="right")

```



## Descriptive measures for Math Score by Socio-economic status


```{r}
group_by(homework, ses) %>%
  summarise(
    count = n(),
    mean = mean(mathsach, na.rm = TRUE),
    sd = sd(mathsach, na.rm = TRUE),
    median = median(mathsach, na.rm = TRUE),
    IQR = IQR(mathsach, na.rm = TRUE)
  )

ggplot(homework, aes(x = ses, y = mathsach, col = factor(ses))) + geom_boxplot(outlier.shape = NA) + geom_jitter(width = 0.2) + theme(legend.position="right")
```

### As we have three level categorical variable (Socio-economic status) as independent variable and math and english score as dependent variable. Hence we will use one way anova to test the significant differneces among the groups.

# Assumptions of ANOVA test

## Assumption 1: Normality


```{r}

shapiro.test(homework$engach)

shapiro.test(homework$mathsach)

```


### Assumption 2: Equal Variance

```{r}
bartlett.test(engach~ses , data=homework)


```

### Assumption 3: Independence of sample

#### as we know our sample observations are independent to each other.


## ANOVA Test 

```{r}
summary(aov(engach~factor(ses),data = homework))
```




## Kruskal wallis test for non-parametric group



```{r}
kruskal.test(mathsach ~ ses, data = homework)

```

